Cost- and Deadline-Constrained Provisioning for Scientific Workflow Ensembles in IaaS Clouds

### 2012-04-28 paper/paper.tex
Edit results

In order to account for the user's priority, we used an exponential scoring where the score for an ensemble $e$ is:
%
$$
Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
$$
%
where $Completed(e)$ is the set of workflows in ensemble $e$ that was completed by the algorithm, and $Priority(w)$ is the priority of workflow $w$ such that the highest-priority workflow has $Priority(w)=0$, the next highest workflow has $Priority(w)=1$, and so on. This exponential scoring function gives the highest priority workflow a score that is higher than all the lower-priority workflows combined:
%
$$
2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i}
$$
%
This scoring is consistent with our definition of the problem, which is to complete as many workflows as possible, according to their priorities, given a set budget and deadline.

@@

\TODO{Update the description of the experiment in this paragraph if necessary}

This experiment used a single application, Montage, and a single ensemble
type, uniform unsorted. This application and ensemble type was chosen
arbitrarily for simplicity. It is expected that the application and ensemble
type have little impact on how well the algorithms are able to stay within
constraints.

@@

Fig.~\ref{fig:variances} shows the results of simulations for estimate
errors ranging from $0\%$ to $50\%$. These simulations were done for 10
ensembles of 50 workflows using 10 budgets and 10 deadlines for each error
value and algorithm (3000 simulations per error value). The figure shows box
plots of the ratio of actual cost to budget and actual makespan to deadline.
The actual cost and makespan are the simulated cost and makespan of the
ensemble. The ratio of actual values to constraints indicates whether the
actual value exeeded the constraint: Values less than 1 indicate that the
constraint was not exceeded, and values greater than 1 indicate that the
constraint was exceeded.

Fig.~\ref{fig:variances}.a shows the ratio of actual cost to budget. This
figure illustrates two important algorithm characteristics. First, the dynamic
algorithms very rarely exceeded the budget, even with very large errors. This
indicates that the dynamic algorithms are able to adapt to uncertainties at
runtime to ensure that the constraints are not exceeded, even when the quality
of information available to them is low. Second, unlike the dynamic
algorithms, the static algorithm frequently exceeded the budget constraint; by
large amounts in some cases. This is a result of the fact that the static
algorithm makes all of its decisions before execution and is not able to adapt
to changing circumstances.

@@

Fig.~\ref{fig:variances}.b shows the ratio of actual makespan to budget. The
interesting thing to notice about this figure is that, for all three
algorithms, the deadline constraint is rarely exceeded, even in cases with
very poor quality estimates. For the dynamic algorithms this is a result of
the fact that they can adapt to poor estimates. For the static algorithm this
is a result of the way that SPSS schedules workflows, and not of a
particularly clever optimization. The SPSS algorithm tends to schedule
workflows early, using up the budget long before the deadline is reached. This
is a consequence of the deadline distribution function in SPSS, which prevents
workflows from starting late. This is shown by the gantt chart in
Fig.~\ref{fig:algorithm-example}.b, which illustrates how SPSS tends to pile
up workflows at the beginning of the timeline. In comparison, the dynamic
algorithms tend to spread out workflow starts over the entire duration as
shown in Fig.~\ref{fig:algorithm-example}.a. The consequence of this
behavior is that SPSS plans tend to use up the budget but leave plenty of time
before the deadline. So when the runtime of the plan is increased by
introducing errors, the SPSS plan has some room to expand without exceeding
the deadline.

@@

\TODO{Update the following paragraph depending on what results Maciek included}

Fig.~\ref{fig:delays} shows the ratios of actual values to constraints when
the provisioning delay is increased from 0 seconds up to 15 minutes. These
simulations were done using 10 uniform unsorted ensembles of 50 workflows from
the Montage application, using 10 deadlines (1000 simulations per algorithm
and provisioning delay). As in Fig.~\ref{fig:variances}, the y-axis in each
plot represents the ratio of the actual simulated value to the constraint
value for the budget constraint (actual cost) or the deadline constraint
(actual makespan). These ratios indicate how frequently and by how much the
actual simulated values exceeded (or did not exceed) the constraints.

@@

\TODO{Update this next paragraph depending on what Maciek included in the figure. The results seem to have changed.}

The effect of provisioning delays on workflow performance is similar to that of inaccurate runtime estimates: when the delays are small, all algorithms are able to produce results within the constraints, but when the delays are larger, the dynamic algorithms are able to adapt to avoid exceeding the constraints while the static algorithm is not. The explanation for this behavior is also similar. The difference is that, a very small error in runtime estimates has little impact on the budget, while a very small provisioning delay has a large impact on the budget. Fig.~\ref{fig:delays}.a shows that even for a very short provisioning delay of 30 seconds (usually it takes more than than 30 seconds just to boot a VM), the SPSS algorithm exceeds the budget by as much as a factor of 2, and has a distribution with an upper quartile of 1.5 times the budget. In the case of delays more than 5 minutes, which is more typical of what has been observed on academic infrastructure clouds~\cite{Juve2011} such as Magellan~\cite{Magellan} and FutureGrid~\cite{FutureGrid}, fully three quarters of the simulations exceeded the budget. With the exception of a few outliers for 10 and 15 minute delays, the dynamic algorithms consistently met the budget and deadline constraints.

@@

\TODO{Update this paragraph when Maciek explains what was included in the plot.}

This behavior suggests that SPSS is far too sensitive to provisioning delays in its current form to be of practical use in real systems. It is likely, as was the case for inaccurate runtime estimates, that modifying the SPSS algorithm would improve its performance for provisioning delays. In fact, since provisioning operations are infrequent (because all the algorithms tend to provision resources for a long time), it is likely that the performance of SPSS could be improved significantly by simply adding an estimate of the provisioning delay to its scheduling function. Such an estimate may not have to be particularly accurate to get good results, and developing an estimate from historical data should be relatively simple. Testing this idea is left for future work.

@@

%This problem differs from previous work on grid and utility grid scheduling in that cloud infrastructures provide more control over provisioning, so that the number of resources can be adjusted according to the requirements of the application. Therefore the problem space becomes larger; it requires not only an efficient mapping of tasks to available resources, but also the selection of the best resource provisioning plan.

%Formulating the problem as a maximization of the number of prioritized workflows completed from the ensemble is also novel and requires workflows to be admitted or rejected based on their estimated resource demands. We believe that this bi-constrained problem is highly relevant because such constraints are commonly imposed on many real-world projects. The approach is also directly applicable to grid environments that provide resource reservations and charge service units for resource use.

@@

%In addition, we found that SPSS tends to perform better on coarse-grained workflows than the dynamic algorithms. For wide, fine-grained workflows, such as CyberShake and Montage, however, the dynamic algorithms frequently produce performance that is as good or better than SPSS, because they are better able to pack the smaller tasks onto idle VMs close to the deadline than SPSS, which distributes deadlines to tasks in a way that prevents them from starting late.

%For very large workflows and ensembles, we found that the planning time of the SPSS algorithm is prohibitive. SPSS often took 10 minutes or more to plan ensembles of 100 workflows with 800 tasks each. This suggests that ensembles of workflows with tens of thousands of tasks, which are commonly encountered in real workflow applications, would take many hours to plan using SPSS. This makes the dynamic algorithms much more attractive for large scale problems.

@@

%The results of this study can be applied to develop tools that assist researchers in planning their large-scale computational experiments. The estimates of cost, runtime, and number of workflows completed that can be obtained from both the static algorithms and from the simulation runs, constitute valuable hints for planning ensembles and evaluating the associated trade-offs.

@@

% This balances the columns by inserting a column break at the given reference
\IEEEtriggeratref{19}

>>>>

In order to account for the user's priority, we used an exponential scoring
where the score for an ensemble $e$ is:
%
$$
Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
$$
%
where $Completed(e)$ is the set of workflows in ensemble $e$ that was
completed by the algorithm, and $Priority(w)$ is the priority of workflow $w$
such that the highest-priority workflow has $Priority(w)=0$, the next highest
workflow has $Priority(w)=1$, and so on. This exponential scoring function
gives the highest priority workflow a score that is higher than all the lower-
priority workflows combined:
%
$$
2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i}
$$
%
This scoring is consistent with our definition of the problem, which is to
complete as many workflows as possible, according to their priorities, given a
set budget and deadline.

@@

@@

Fig.~\ref{fig:variances} shows the results of simulations for estimate errors
ranging from $0\%$ to $50\%$. The figure summarizes the results of an
extensive suite of 525,000 simulations (10 ensembles of 50 workflows x 5
applications x 5 distributions x 10 budgets x 10 deadlines x 7 error values x
3 algorithms). The figure shows box plots of the ratio of actual cost to
budget and actual makespan to deadline. The actual cost and makespan are the
simulated cost and makespan of the ensemble. The ratio of actual values to
constraints indicates whether the actual value exeeded the constraint: Values
greater than 1 indicate that the constraint was exceeded.

Fig.~\ref{fig:variances}.a shows the ratio of actual cost to budget. This
plot illustrates two important algorithm characteristics. First, the dynamic
algorithms very rarely exceeded the budget, even with very large errors. This
indicates that the dynamic algorithms are able to adapt to uncertainties at
runtime to ensure that the constraints are not exceeded, even when the quality
of information available to them is low. Second, unlike the dynamic
algorithms, the static algorithm frequently exceeded the budget constraint; by
large amounts in some cases. This is a result of the fact that the static
algorithm makes all of its decisions before execution and is not able to adapt
to changing circumstances.

@@

Fig.~\ref{fig:variances}.b shows the ratio of actual makespan to budget. The
interesting thing to notice about this plot is that, for all three
algorithms, the deadline constraint is rarely exceeded, even in cases with
very poor quality estimates. For the dynamic algorithms this is a result of
the fact that they can adapt to poor estimates. For the static algorithm this
is a result of the way that SPSS schedules workflows, and not of a
particularly clever optimization. The SPSS algorithm tends to schedule
workflows early, using up the budget long before the deadline is reached. This
is a consequence of the deadline distribution function in SPSS, which prevents
workflows from starting late. This is shown by the gantt chart in
Fig.~\ref{fig:algorithm-example}.b, which illustrates how SPSS tends to pile
up workflows at the beginning of the timeline. In comparison, the dynamic
algorithms tend to spread out workflow starts over the entire duration as
shown in Fig.~\ref{fig:algorithm-example}.a. The consequence of this
behavior is that SPSS plans tend to use up the budget but leave plenty of time
before the deadline. So when the runtime of the plan is increased by
introducing errors, the SPSS plan has some room to expand without exceeding
the deadline.

@@

Fig.~\ref{fig:delays} shows the ratios of actual values to constraints when
the provisioning delay is increased from 0 seconds up to 15 minutes. The
figure summarizes a suite of 105,000 simulations (10 ensembles of 50 workflows
x 5 distributions x 10 budgets x 10 deadlines x 7 error values x 3
algorithms). Only one application, Montage, was used in this experiment. As in
Fig.~\ref{fig:variances}, the y-axis in each plot represents the ratio of the
actual simulated value to the constraint value for the budget constraint
(actual cost) or the deadline constraint (actual makespan). These ratios
indicate how frequently and by how much the actual simulated values exceeded
(or did not exceed) the constraints.

@@

The effect of provisioning delays on workflow performance is similar to that
of inaccurate runtime estimates: when the delays are small, all algorithms are
able to produce results within the constraints, but when the delays are
larger, the dynamic algorithms are able to adapt to avoid exceeding the
constraints while the static algorithm is not. In the case of delays of more
than 5 minutes, which is typical of what has been observed on academic
infrastructure clouds~\cite{Juve2011} such as Magellan~\cite{Magellan} and
FutureGrid~\cite{FutureGrid}, approximately half of the SPSS simulations
exceeded the budget while none of the simulations using a dynamic algorithm
exceeded the budget.

@@

This experiment suggests that SPSS is far too sensitive to provisioning delays
in its current form to be of practical use in real systems. It is possible
that modifying the SPSS algorithm to account for provisioning delays would
improve its performance on this experiment. In fact, since provisioning
operations are infrequent (because all the algorithms tend to provision
resources for a long time), it is likely that the performance of SPSS could be
improved significantly by simply adding an estimate of the provisioning delay
to its scheduling function. Such an estimate may not have to be particularly
accurate to get good results, and developing an estimate from historical data
should be relatively simple. Testing this idea is left for future work.

@@

%This problem differs from previous work on grid and utility grid scheduling
%in that cloud infrastructures provide more control over provisioning, so that
%the number of resources can be adjusted according to the requirements of the
%application. Therefore the problem space becomes larger; it requires not only
%an efficient mapping of tasks to available resources, but also the selection
%of the best resource provisioning plan.

%Formulating the problem as a maximization of the number of prioritized
%workflows completed from the ensemble is also novel and requires workflows to
%be admitted or rejected based on their estimated resource demands. We believe
%that this bi-constrained problem is highly relevant because such constraints
%are commonly imposed on many real-world projects. The approach is also
%directly applicable to grid environments that provide resource reservations
%and charge service units for resource use.

@@

%In addition, we found that SPSS tends to perform better on coarse-grained
%workflows than the dynamic algorithms. For wide, fine-grained workflows, such
%as CyberShake and Montage, however, the dynamic algorithms frequently produce
%performance that is as good or better than SPSS, because they are better able
%to pack the smaller tasks onto idle VMs close to the deadline than SPSS,
%which distributes deadlines to tasks in a way that prevents them from
%starting late.

%For very large workflows and ensembles, we found that the planning time of
%the SPSS algorithm is prohibitive. SPSS often took 10 minutes or more to plan
%ensembles of 100 workflows with 800 tasks each. This suggests that ensembles
%of workflows with tens of thousands of tasks, which are commonly encountered
%in real workflow applications, would take many hours to plan using SPSS. This
%makes the dynamic algorithms much more attractive for large scale problems.

@@

%The results of this study can be applied to develop tools that assist
%researchers in planning their large-scale computational experiments. The
%estimates of cost, runtime, and number of workflows completed that can be
%obtained from both the static algorithms and from the simulation runs,
%constitute valuable hints for planning ensembles and evaluating the
%associated trade-offs.

@@

% This balances the columns by inserting a column break at the given reference
\IEEEtriggeratref{16}
